LME-MPM applied to quasi-brittle fracture.

### 2020-05-29 _region_.tex
Cambios 29-5-2020

opening $w_n$ within the band width, $h_{\epsilon}$. Therefore, 

@@

\subsection{$\epsilon$-neighborhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
\PNA{The construction of the $\epsilon$-neighborhood is a major issue of the proposed methodology.} Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}, \PNA{the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error} prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global definition of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. \DM{Intead, a} linked
list of tributary nodes is reconstructed with the closest node as
starting point at each time step \DM{deleted instead}. Therefore, each node is assigned to an
unique linked-list of those particles which are closest to it. This
approach allows to reconstruct easily a search list with the support
of the initial connectivity of the mesh and offers a suitable basis
for an easy implementation of other interpolation techniques with
minimum coding effort. A detailed explanation of the proposed
algorithm can be found in \ref{sec:node-linked-method-1}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cases of study and discussion}
\label{sec:3}

\DM{The proposed approach to overcome the two main shortcomings of EighenMPM \cite{Zhang_EE_2020} has been evaluated with "two benchmark tests"¿?. } \DM{DELETE: As noticed in \cite{Zhang_EE_2020}, the EigenMPM have two important
shortcomings}. The first one deals with the presence of stress instabilities provoked by the spatial discretization, appearing even
when the \acrshort{gimp} shape functions are employed. To overcome it, \acrshort{lme}
approximants are introduced as an
alternative to the existing interpolation techniques. Its
performance in mitigating spurious stress oscillations under the \acrshort{mpm}~\cite{Wobbes2020,Molinos2020} framework helps to enhance
notoriously the quality of the results as it is observed in Section
\ref{sec:3.1}, where both interpolation techniques
are compared by carrying out an eigenerosion simulation. The second
limitation is concerned to its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}.\MMP{A proof
of it is exposed in Sections \ref{sec:3.2} \ref{sec:3.3}, where experimental results
are compared with experimental ones as well as eigensoftening
computations. In the first case comparing with a semicircular bending
test and in the second case with a drop-weight impact test.}

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation. The application
consists of a square plate of size $H = 1$ containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\PNA{UNIDADES}

@@

The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \gls{E} = 1.06, the
Poisson's ratio \gls{nu} = 0.333, and critical energy-release rate
\gls{Gf} = 0.0001. This set of parameters has been previously employed by Pandolfi~\textit{et al.}~\cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretization are required. On the one hand,
a cartesian grid of nodes is considered with a nodal spacing value of
0.025. On the other hand, the plate will be modeled with a initial
layout of four particles per element and occupying the Gauss quadrature
positions.\\

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in the reaction-displacements curve, in both loading and post-failure stages,
when linear interpolation technique is used. In contrast, \acrshort{lme} simulation does not
exhibit these spurious oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the differences in the obtained peaks are not extremely significant, it can be owing to the aforementioned hookean material employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive models will be employed, severe
%if hyperelastic materials are consider for future simulations,
inaccuracies may appear in the stress field that could affect dramatically to the
final result. Focusing on the post failure behavior, it can be seen how \acrshort{lme} produces soft stress
field evolution even after breaking.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the
  energy is release dinamically by the crack, the waves propagate
  along the domain.}

@@

% The second application to be considered is a series of semicircular
% bending tests on Johnstone, which have been carried out by Lim et al.95 The aim of this example is to examine its predictive capability of the SPH model under a wide range
% of mixed-mode loading conditions. The geometry and boundary conditions of the test in the experiment are shown in
% Figure 9A. Similar with 3-point bending test, the semicircular specimens with a radius of 47.5 mm and a thickness of
% 20 mm are supported by two rollers at a span of 47.5 mm and loaded by another roller on its mid-span. Notches with a
% length of 16.6 mm are created with different notch angles varying from 0 to 60 with respect to the vertical axis to
% investigate the influences of the notch angles on the peak load (P) and the fracture toughness (KI,KII), which are calcu-
% lated as follows:

% where KI and KII are respectively mode I and mode II fracture toughness; r is the specimen radius; t is the specimen
% thickness; a is the notch length; YI and YII are mode I and mode II normalised stress intensity factor. As the values of YI
% and YII can be obtained from the data by Lim at al.6 KI and KII can be calculated based on the above equation once the
% peak load (P) is known. The material properties of Johnstone are listed in Table 1. The friction angle and dilation angle
% are respectively taken as 33 and 4  for average quality rock masses in Hoek and Brown96 for the following examples.
% Figure 9 shows the corresponding SPH configuration of the semicircular bending test with a notch angle of 30 . In
% the tests, a particle discretisation of 0.95 mm is used, resulting in a total number of 3995 SPH particles for generating
% the specimen. To create accurate notch angles using reasonable particle numbers, the initial SPH arrangement in this
% test is different from the traditional one such as the 3-point bending test in Section 5.1. The initial location of each SPH
% particle in semicircular bending tests is related to the notch angle as shown in Figure 9B. Similar to the 3-point bending
% tests, three layers of boundary particles are generated for simulating the point loading in the experiments. The velocities
% of the top boundary particles are vx = 0,vy = − 2 mm/s, whereas the velocities of the bottom boundary particles are
% as vy = 0. As Johnstone rocks are classified as soft rocks with a relatively small elastic Young's modulus, both artificial
% viscosity and artificial stress are used in this test to stabilise the numerical solutions. The ramp loading is also applied
% to alleviate the dynamic effect due to the sudden loading. The stresses of boundary particles are updated in the same
% way as that in the 3-point bending test. The stresses of real particles near the loading particles are recorded during the
% simulation process, from which the peak load and fracture toughness can be obtained for the comparison with the
% experimental data.
% Figure 10 illustrates the progressive development of fracture and the horizontal velocity profile at the failure stage
% for the semicircular specimen with the rotational angle of 30 . As seen in Figure 10A, some particles near the tip of the
% notch are first damaged, showing that the fracture is initiated at the location of the notch tip. Then, the fracture propa-
% gates towards the loading point (Figure 10B–D), and eventually, a vertical curvilinear fracture pattern is formed in the
% middle of the sample (Figure 10E). The horizontal velocity profile corresponding to Figure 10E is also plotted in
% Figure 10F. It can be seen that the particles on the left side of the fracture move to the left whereas the particles on the
% other side move to the right direction; thus, the sample is split into two halves at the end. In this approach, SPH particle
% itself could not be cracked visually; however, each damaged SPH particle possesses a FPZ, and the material behaviour
% within the FPZ will contribute to the decrease in the overall stress state of the SPH particle, thereby resulting in the sep-
% aration between two damaged SPH particles. The same test without using the artificial stress is also plotted in
% Figure 11. The result remains stable when the fracture is initiated (Figure 11A) but becomes unstable during the post-
% failure process (Figure 11B,C). Unrealistic fracture and particles clustering are observed in Figure 11B,C, which indi-
% cates the tensile instability problem is occurred in SPH. By comparing those two results, it can be concluded that the

% tensile instability problem can be completely removed and good numerical stability can be achieved by just adopting
% this artificial stress term.
% Figure 12 depicts the comparison of mixed-mode fracture envelope between experiments and SPH simulations for
% different notch inclination angles α = 0 , 15  , 30  , 45 , 50 and 60 . It shows that the simulation fracture patterns
% marked as red lines match well with the experimental results shown as black lines. The final fracture pattern is a verti-
% cal straight line when α = 0 (pure mode I). When α > 0 , the sample is subjected to mixed-mode or pure mode II load-
% ing, resulting in a curvilinear fracture paths. For all simulation cases, the fractures initiate from the notch tip and then
% propagate towards upper loading point, while the experiments show that a few fractures tend to initiate behind the
% notch tip when the notch angle is beyond α =50 . Nevertheless, the simulation results still could successfully predict
% the overall fracture envelope in experiments and demonstrate the
% capability of the traditional SPH model for capturing

% the fracture development under various complex loading conditions. In the meanwhile, the comparison of peak load
% between experiments and simulations for the notch inclination angles varying from 0 to 60 with an internal incre-
% ment of 5 is plotted in Figure 13A. It is noted here that a notch angle of 54 (instead of 55) is used in the current test
% as the sample is subjected to pure mode II loading condition when α = 54 .95 The SPH simulation can well capture the
% increasing trend of the peak load with the increasing notch angles in the experiment. Additionally, Figure 13B com-
% pares the predictions of fracture toughness in SPH simulation against the experimental results and Lim's theory.95 The
% simulation results perfectly fall within the experimental data ranges even though the fracture toughness data in the
% experiments are quite scattered. Nevertheless, the numerical results agree closely with the empirical fracture toughness
% envelope, which was proposed by Lim et al.95 and represented as a black line Figure 13B. It further confirms that our
% model could accurately predict the mixed-mode fracture behaviour of rocks in semicircular bending test.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    1540    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\cite{Wang_2020}

@@

\end{document}
\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-1244) }

>>>>

opening $w_n$ within the band width $h_{\epsilon}$. Therefore, 

@@

\subsection{$\epsilon$-neighbourhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
\PNA{The construction of the $\epsilon$-neighbourhood is a major issue of the proposed methodology.} Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}, \PNA{the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error} prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global support of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. Instead, a linked
list with the surrounding nodes for each node is defined at the initial
time. Also, each particle has to be associated with the
closest node instead with the belonging element as heretofore has been
done.
\MMP{Now lets define the concept of a particle \textit{tributary nodes} as the
list of nodes close to it among which the information transfers
occurs. Once the closest node are defined, the particle
posses automatically a list of \textit{tributary nodes}. This has two
immediate benefits, the first is that shape functions with global
support as \acrshort{lme} are easier to implement,
and the second is the increase of the efficiency of the $\epsilon$-neighbourhood
reconstruction. Regarding this last benefit, since the current
particle has a \textit{tributary nodes} list assigned to it, and
consequently each node has a linked list with a set of
\textit{tributary particles}\footnote{Notice that the list of the
  particles close to each node is unique for each node and is updated
  in each time step.} close to it, is possible to define efficiently a
search list with those particles close to the current particle.
In the end, particles out of the area of influence, Figure
\ref{fig:Failed-particles}, are pop out of the final
$\epsilon$-neighbourhood. A final remark about this new approach is
its ability to reconstruct easily a search list with the support of
the initial connectivity of the mesh. It provides a suitable basis for
an easy implementation of other interpolation techniques with minimum
coding effort. A detailed explanation of the proposed algorithm can be
found in \ref{sec:node-linked-method-1}.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Cases of study and discussion}
\label{sec:3}

\DM{The proposed approach to overcome the two main shortcomings of
  EighenMPM \cite{Zhang_EE_2020} has been evaluated with ``three
  benchmark tests''}.
The first one deals with the presence of stress
instabilities provoked by the spatial discretization, appearing even when the \acrshort{gimp} shape functions are employed. To overcome it, \acrshort{lme}
approximants are introduced as an alternative to the existing interpolation techniques. Its
performance in mitigating spurious stress oscillations under the \acrshort{mpm}~\cite{Wobbes2020,Molinos2020} framework helps to enhance
notoriously the quality of the results as it is observed in Section
\ref{sec:3.1}, where both interpolation techniques
are compared by carrying out an eigenerosion simulation. The second
limitation is concerned to its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}.\MMP{A proof
of it is exposed in Sections \ref{sec:3.2} \ref{sec:3.3}, where experimental results
are compared with experimental ones as well as eigensoftening
computations. In the first case comparing with a semicircular bending
test and in the second case with a drop-weight impact test.}

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation. The application
consists of a square plate of size\MMP{$H = 1\ [L] $} containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \MMP{\gls{E} = 1.06
$[M] / ([L] \cdot [T]^2)$, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001 $[M] / \cdot
[T]^2$. The absence of units is due to the current simulation will not be validated against experimental results, where
scale factor is relevant, but will be against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}.} In
\acrshort{mpm}, two different discretization are required. On the one hand, a cartesian grid of
nodes is considered with a nodal spacing value of 0.025. On the other
hand, the plate will be modeled with a initial layout of four
particles per element and occupying the Gauss quadrature positions.\\

@@

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in the reaction-displacements curve, in both
loading and post-failure stages, when linear interpolation technique
is used. In contrast, \acrshort{lme} simulation does not exhibit these
spurious oscillations and remains linear until the failure. About the
strength, linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the
  differences in the obtained peaks are not extremely significant, it
  can be owing to the aforementioned hookean material
  employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive
  models will be employed, severe inaccuracies may appear in the
  stress field that could affect dramatically to the final 
  result. Focusing on the post failure behavior, it can be seen how
  \acrshort{lme} produces soft stress field evolution even after
  breaking.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the
  energy is release dinamically by the crack, the waves propagate
  along the domain.}

@@

The second application to be considered is two semicircular
bending tests on Johnstone, which have been carried out by Lim et
al.95. Wang {\it et al.}\cite{Wang_2020} proposed it with the aim of
examine its predictive capability of the \arcshort{sph} model under a
wide range of mixed-mode loading conditions. Nonetheless, here only
both extreme cases 0$^º$ and 60$^º$ will be reproduced.

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test} . Similar with 3-point bending test,
the semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm are supported by two rollers at a span of 47.5 mm and loaded by
another roller on its mid-span. Notches with a length of 16.6 mm are
created with different notch angles varying from 0 to 60 with
respect to the vertical axis to investigate the influences of the
notch angles on the peak load (P).

@@

The material properties of Johnstone are listed in Table \ref{tab:Johnstone-properties} The friction angle and dilation angle
are respectively taken as 33 and 4  for average quality rock
masses in Hoek and Brown96 for the following examples.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    1540    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Figure 9 shows the corresponding SPH configuration of the semicircular
bending test with a notch angle of 30 . In the tests, a particle
discretisation of 0.95 mm is used, resulting in a total number of 3995
SPH particles for generating the specimen. To create accurate notch
angles using reasonable particle numbers, the initial SPH arrangement
in this test is different from the traditional one such as the 3-point
bending test in Section 5.1. The initial location of each SPH particle in semicircular bending tests is related to the notch angle as shown in Figure 9B. Similar to the 3-point bending
tests, three layers of boundary particles are generated for simulating
the point loading in the experiments. The velocities of the top boundary particles are vx = 0,vy = − 2 mm/s, whereas the velocities of the bottom boundary particles are
as vy = 0.

As Johnstone rocks are classified as soft rocks with a relatively
small elastic Young's modulus, both artificial viscosity and
artificial stress are used in this test to stabilise the numerical
solutions. The ramp loading is also applied to alleviate the dynamic
effect due to the sudden loading.

The stresses of boundary particles are updated in the same way as that
in the 3-point bending test. The stresses of real particles near the
loading particles are recorded during the simulation process, from
which the peak load and fracture toughness can be obtained for the
comparison with the experimental data.

Figure 10 illustrates the progressive development of fracture and the
horizontal velocity profile at the failure stage for the semicircular specimen with the rotational angle of 30 . As seen in Figure 10A, some particles near the tip of the
notch are first damaged, showing that the fracture is initiated at the
location of the notch tip. Then, the fracture propagates towards the
loading point (Figure 10B–D), and eventually, a vertical curvilinear
fracture pattern is formed in the middle of the sample (Figure
10E). The horizontal velocity profile corresponding to Figure 10E is
also plotted in Figure 10F. It can be seen that the particles on the
left side of the fracture move to the left whereas the particles on
the other side move to the right direction; thus, the sample is split
into two halves at the end. In this approach, SPH particle itself could not be cracked visually; however, each damaged SPH particle possesses a FPZ, and the material behaviour
within the FPZ will contribute to the decrease in the overall stress
state of the SPH particle, thereby resulting in the separation between
two damaged SPH particles. The same test without using the artificial
stress is also plotted in Figure 11. The result remains stable when
the fracture is initiated (Figure 11A) but becomes unstable during the
post-failure process (Figure 11B,C). Unrealistic fracture and
particles clustering are observed in Figure 11B,C, which indicates the
tensile instability problem is occurred in SPH. By comparing those two
results, it can be concluded that the tensile instability problem can
be completely removed and good numerical stability can be achieved by
just adopting this artificial stress term. Figure 12 depicts the
comparison of mixed-mode fracture envelope between experiments and SPH
simulations for different notch inclination angles α = 0 , 15  ,
30  , 45 , 50 and 60 . It shows that the simulation fracture
patterns marked as red lines match well with the experimental results
shown as black lines. The final fracture pattern is a vertical
straight line when α = 0 (pure mode I). When α > 0 , the sample is
subjected to mixed-mode or pure mode II loading, resulting in a curvilinear fracture paths. For all simulation cases, the fractures initiate from the notch tip and then
propagate towards upper loading point, while the experiments show that
a few fractures tend to initiate behind the notch tip when the notch angle is beyond α =50 . Nevertheless, the simulation results still could successfully predict
the overall fracture envelope in experiments and demonstrate the
capability of the traditional SPH model for capturing the fracture development under various complex loading conditions. In the meanwhile, the comparison of peak load
between experiments and simulations for the notch inclination angles
varying from 0 to 60 with an internal increment of 5 is plotted
in Figure 13A. It is noted here that a notch angle of 54 (instead of
55) is used in the current test as the sample is subjected to pure
mode II loading condition when α = 54 .95 The SPH simulation can
well capture the increasing trend of the peak load with the increasing
notch angles in the experiment. Additionally, Figure 13B compares the
predictions of fracture toughness in SPH simulation against the
experimental results and Lim's theory.95 The simulation results perfectly fall within the experimental data ranges even though the fracture toughness data in the
experiments are quite scattered. Nevertheless, the numerical results
agree closely with the empirical fracture toughness envelope, which
was proposed by Lim et al.95 and represented as a black line Figure
13B. It further confirms that our model could accurately predict the
mixed-mode fracture behaviour of rocks in semicircular bending test.

@@

\end{document}

\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-1283) }
